Spotify Analysis by Don Nguyen
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
raw <- read.csv("spotify_data.csv", sep = '\t')

drops <- c("X","analysis_url", "id", "track_href", "type", "uri")
songs <- raw[ , !(names(raw) %in% drops)]
```

# Univariate Plots Section
```{r echo=FALSE, Univariate_Summary}
summary(subset(songs, songs$like.=='yes'))
summary(subset(songs, songs$like.=='no'))
```

```{r echo=FALSE, Univariate_Plots}
ggplot(aes(x=acousticness), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Acousticness")
ggplot(aes(x=danceability), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Danceability")
ggplot(aes(x=energy), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Energy")
ggplot(aes(x=instrumentalness), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Instrumentalness")
ggplot(aes(x=liveness), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Liveness")
ggplot(aes(x=loudness), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Loudness")
ggplot(aes(x=speechiness), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Speechiness")
ggplot(aes(x=tempo), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Tempo")
ggplot(aes(x=valence), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Valence")
```

# Univariate Analysis

### What is the structure of your dataset?
The dataset consists of 2000 songs, half of which are songs I categorized as liking on Spotify, the other half categorized as not liking. The songs I do not like have been pulled from many different genres of music. Each song has 14 features that can be analyzed.

### What is/are the main feature(s) of interest in your dataset?
The main feature to be investigated is the "like" column, which represents whether or not I have liked the song on Spotify.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
I will be exploring all other features of the songs as a function of "like" to find patterns in songs I do and do not enjoy.

### Did you create any new variables from existing variables in the dataset?
Currently, none of the variables have been created based off existing variables.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?



# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?




# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
