Spotify Analysis by Don Nguyen
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
raw <- read.csv("spotify_data.csv", sep = '\t')

drops <- c("X","analysis_url", "id", "track_href", "type", "uri")
songs <- raw[ , !(names(raw) %in% drops)]
```

# Univariate Plots Section
```{r echo=FALSE, Univariate_Summary}
summary(subset(songs, songs$like.=='yes'))
summary(subset(songs, songs$like.=='no'))
```

```{r echo=FALSE, Univariate_Plots}
ggplot(aes(x=acousticness), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Acousticness")
ggplot(aes(x=danceability), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Danceability")
ggplot(aes(x=duration_ms), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Duration (ms)")
ggplot(aes(x=energy), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Energy")
ggplot(aes(x=instrumentalness), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Instrumentalness")
ggplot(aes(x=liveness), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Liveness")
ggplot(aes(x=loudness), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Loudness")
ggplot(aes(x=speechiness), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Speechiness")
ggplot(aes(x=tempo), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Tempo")
ggplot(aes(x=valence), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Valence")
```
```{r echo=FALSE, Univariate_Box_Plots}
ggplot(aes(x=like., y=acousticness), data = songs) + geom_boxplot() + ggtitle("Acousticness")
ggplot(aes(x=like., y=danceability), data = songs) + geom_boxplot() + ggtitle("Danceability")
ggplot(aes(x=like., y=duration_ms), data = songs) + geom_boxplot() + ggtitle("Duration (ms)")
ggplot(aes(x=like., y=energy), data = songs) + geom_boxplot() + ggtitle("Energy")
ggplot(aes(x=like., y=instrumentalness), data = songs) + geom_boxplot() + ggtitle("Instrumentalness")
ggplot(aes(x=like., y=liveness), data = songs) + geom_boxplot() + ggtitle("Liveness")
ggplot(aes(x=like., y=loudness), data = songs) + geom_boxplot() + ggtitle("Loudness")
ggplot(aes(x=like., y=speechiness), data = songs) + geom_boxplot() + ggtitle("Speechiness")
ggplot(aes(x=like., y=tempo), data = songs) + geom_boxplot() + ggtitle("Tempo")
ggplot(aes(x=like., y=valence), data = songs) + geom_boxplot() + ggtitle("Valence")
```

# Univariate Analysis

### What is the structure of your dataset?
The dataset consists of 2000 songs, half of which are songs I categorized as liking on Spotify, the other half categorized as not liking. The songs I do not like have been pulled from many different genres of music. Each song has 14 features that can be analyzed.

### What is/are the main feature(s) of interest in your dataset?
The main feature to be investigated is the "like" column, which represents whether or not I have liked the song on Spotify.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
I will be exploring all other features of the songs as a function of "like" to find patterns in songs I do and do not enjoy.

### Did you create any new variables from existing variables in the dataset?
Currently, none of the variables have been created based off existing variables.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
A couple data points needed to be removed prior to loading into R since they had null features. After removing the two songs, the rest of the data came tidy from Spotify. A few features were dropped due to not giving much information to analyze. 

The features that stood out to me as having the most difference between songs I liked and disliked were energy, acousticness, and instrumentalness. Songs I liked tend to have lower energy, and more acousticness than songs I did not like. The songs I dislike also are clearly not instrumental, whereas songs I like have a higher probability of being instrumental.


# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}
ggplot(aes(x=loudness, y=energy), data = songs) + geom_point() + ggtitle("Relationship Between Loudness and Energy")
ggplot(aes(x=loudness, y=valence), data = songs) + geom_point() + ggtitle("Relationship Between Loudness and Valence")
ggplot(aes(x=valence, y=energy), data = songs) + geom_point() + ggtitle("Relationship Between Happy Songs and Song Energy Level")
ggplot(aes(x=liveness, y=energy), data = songs) + geom_point() + ggtitle("Relationship Between Live Songs and Energy")
ggplot(aes(x=acousticness, y=danceability), data = songs) + geom_point() + ggtitle("Danceability of Acoustic Songs")
ggplot(aes(x=instrumentalness, y=danceability), data = songs) + geom_point() + ggtitle("Danceability of Instrumental Songs")
ggplot(aes(x=tempo, y = danceability), data = songs) + geom_point() + ggtitle("Tempo vs. Danceability")
ggplot(aes(x=acousticness, y=liveness), data = songs) + geom_point() + ggtitle("Relationship Between Acousticness and Probaiblity of Liveness")
```
```{r echo=FALSE, Bivariate_Correlations}
with(songs, cor.test(loudness, energy))
with(songs, cor.test(valence, energy))
with(songs, cor.test(tempo, danceability))
```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?
There is a clear positive linear relationship between loudness and energy. There seems to be a polynomial relationship between tempo and danceability too. These relationships are not surprising. However, valence did not show a strong correlation with energy nor loudness. The other variables also did not show any strong correlations with each other.

### What was the strongest relationship you found?
Loudness and energy has a strong, positive linear relationship. Energy may be based off of loudness.



# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
