Spotify Analysis by Don Nguyen
========================================================

This analysis will look at a data set pulled from Spotify API, using Python. The data contains 1000 songs which I prefer and 1000 songs from various genres that are not preferred. The Spotify allows access to song data such as "acousticness", "danceability", and "energy". More information about these features can be found on [Spotify's documentation](https://developer.spotify.com/web-api/get-audio-features/).

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.
#install.packages('rpart')
#install.packages('rpart.plot')
#install.packages('rattle')
#install.packages('randomForest')
library(ggplot2)
library(rattle)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(randomForest)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
raw <- read.csv("spotify_data.csv", sep = '\t')

drops <- c("X","analysis_url", "id", "track_href", "type", "uri")
songs <- raw[ , !(names(raw) %in% drops)]
```

# Univariate Plots Section
The first task is to look at the data from a broad overview. Shown below is a summary of the features in songs that are preferred and songs that are not preferred.

```{r echo=FALSE, Univariate_Summary}
summary(subset(songs, songs$like.=='yes'))
summary(subset(songs, songs$like.=='no'))
```
By plotting histograms of each feature, the distributions of the data can be explored. Each feature's distribution was compared between preferred songs and songs that were not preferred.

```{r echo=FALSE, Univariate_Plots}
ggplot(aes(x=acousticness), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Acousticness")
ggplot(aes(x=danceability), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Danceability")
ggplot(aes(x=duration_ms), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Duration (ms)")
ggplot(aes(x=energy), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Energy")
ggplot(aes(x=instrumentalness), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Instrumentalness")
ggplot(aes(x=liveness), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Liveness")
ggplot(aes(x=loudness), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Loudness")
ggplot(aes(x=speechiness), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Speechiness")
ggplot(aes(x=tempo), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Tempo")
ggplot(aes(x=valence), data = songs) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + facet_wrap(~like.) + ggtitle("Valence")
```

A similar analysis can be done by creating box plots of each variable. Box plots give a better visual comparison of the differences of means between preferred songs and songs that are not preferred.

```{r echo=FALSE, Univariate_Box_Plots}
ggplot(aes(x=like., y=acousticness), data = songs) + geom_boxplot() + ggtitle("Acousticness")
ggplot(aes(x=like., y=danceability), data = songs) + geom_boxplot() + ggtitle("Danceability")
ggplot(aes(x=like., y=duration_ms), data = songs) + geom_boxplot() + ggtitle("Duration (ms)")
ggplot(aes(x=like., y=energy), data = songs) + geom_boxplot() + ggtitle("Energy")
ggplot(aes(x=like., y=instrumentalness), data = songs) + geom_boxplot() + ggtitle("Instrumentalness")
ggplot(aes(x=like., y=liveness), data = songs) + geom_boxplot() + ggtitle("Liveness")
ggplot(aes(x=like., y=loudness), data = songs) + geom_boxplot() + ggtitle("Loudness")
ggplot(aes(x=like., y=speechiness), data = songs) + geom_boxplot() + ggtitle("Speechiness")
ggplot(aes(x=like., y=tempo), data = songs) + geom_boxplot() + ggtitle("Tempo")
ggplot(aes(x=like., y=valence), data = songs) + geom_boxplot() + ggtitle("Valence")
```

Several t-tests can be computed for the features to quantify the difference in means, assuming the data follows a normal distribution.

```{r echo=FALSE, t_Tests}
like = subset(songs, like.=='yes')
dislike = subset(songs, like.=='no')
t.test(like$instrumentalness, dislike$instrumentalness,alternative="two.sided",mu=0,var.equal=F,conf.level=0.99)
t.test(like$energy, dislike$energy,alternative="two.sided",mu=0,var.equal=F,conf.level=0.99)
t.test(like$acousticness, dislike$acousticness,alternative="two.sided",mu=0,var.equal=F,conf.level=0.99)
t.test(like$danceability, dislike$danceability,alternative="two.sided",mu=0,var.equal=F,conf.level=0.99)
t.test(like$tempo, dislike$tempo,alternative="two.sided",mu=0,var.equal=F,conf.level=0.99)
t.test(like$valence, dislike$valence,alternative="two.sided",mu=0,var.equal=F,conf.level=0.99)
```

# Univariate Analysis

### What is the structure of your dataset?
The dataset consists of 2000 songs, half of which are songs I categorized as preferring on Spotify, the other half categorized as not preferring. The songs I do not prefer have been pulled from many different genres of music. Each song has 14 features that can be analyzed.

### What is/are the main feature(s) of interest in your dataset?
The main feature to be investigated is the "like." column, which represents whether or not I have classified the song preferrence on Spotify.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
I will be exploring all other features of the songs as a function of "like." to find patterns in songs I do and do not enjoy. The features that stand out initally are energy and instrumentalness.

### Did you create any new variables from existing variables in the dataset?
Currently, none of the variables have been created based off existing variables.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
A couple data points needed to be removed prior to loading into R since they had null features. After removing the two songs, the rest of the data came tidy from Spotify. A few features were dropped due to not giving much information to analyze. 

The features that stood out to me as having the most difference between songs I liked and disliked were energy, acousticness, and instrumentalness. Songs I prefer tend to have lower energy, and more acousticness than songs I did not prefer. The songs I do not prefer also are clearly not instrumental, whereas songs I prefer have a higher probability of being instrumental.


# Bivariate Plots Section
Scatter plots are created to find and understand relationships between the features.

```{r echo=FALSE, warning=FALSE, message=FALSE, Bivariate_Plots}
ggplot(aes(x=loudness, y=energy), data = subset(songs, songs$loudness < quantile(songs$loudness, .99) & songs$loudness > quantile(songs$loudness, .01))) + geom_point() + ggtitle("Relationship Between Loudness and Energy") + geom_smooth() 
ggplot(aes(x=loudness, y=valence), data = songs) + geom_point() + ggtitle("Relationship Between Loudness and Valence")
ggplot(aes(x=valence, y=energy), data = songs) + geom_point() + ggtitle("Relationship Between Happy Songs and Song Energy Level")
ggplot(aes(x=liveness, y=energy), data = songs) + geom_point() + ggtitle("Relationship Between Live Songs and Energy")
ggplot(aes(x=acousticness, y=danceability), data = songs) + geom_point() + ggtitle("Danceability of Acoustic Songs")
ggplot(aes(x=instrumentalness, y=danceability), data = songs) + geom_point() + ggtitle("Danceability of Instrumental Songs")
ggplot(aes(x=tempo, y = danceability), data = songs) + geom_point() + ggtitle("Tempo vs. Danceability")
ggplot(aes(x=acousticness, y=liveness), data = songs) + geom_point() + ggtitle("Relationship Between Acousticness and Probaiblity of Liveness")
```

The Pearson's correlation can be used to quantify trends between features.

```{r echo=FALSE, Bivariate_Correlations}
with(songs, cor.test(loudness, energy))
with(songs, cor.test(valence, energy))
with(songs, cor.test(tempo, danceability))
```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?
There is a clear positive linear relationship between loudness and energy. There seems to be a polynomial relationship between tempo and danceability too. These relationships are not surprising. However, valence did not show a strong correlation with energy nor loudness. The other variables also did not show any strong correlations with each other.

### What was the strongest relationship you found?
Loudness and energy has a strong, positive linear relationship. Energy may be based off of loudness.


# Multivariate Plots Section
Plots similar to those in the bivariate section can be filtered by song preferrence to visualize any differences.

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}
ggplot(aes(x=loudness, y=energy), data = subset(songs, songs$loudness < quantile(songs$loudness, .99) & songs$loudness > quantile(songs$loudness, .01))) + geom_point(aes(color = like.), alpha = 1/2) + ggtitle("Relationship Between Loudness and Energy") 
ggplot(aes(x=loudness, y=valence), data = songs) + geom_point(aes(color = like.), alpha = 1/2) + ggtitle("Relationship Between Loudness and Valence")
ggplot(aes(x=valence, y=energy), data = songs) + geom_point(aes(color = like.), alpha = 1/2) + ggtitle("Relationship Between Happy Songs and Song Energy Level")
ggplot(aes(x=liveness, y=energy), data = songs) + geom_point(aes(color = like.), alpha = 1/2) + ggtitle("Relationship Between Live Songs and Energy")
ggplot(aes(x=acousticness, y=danceability), data = songs) + geom_point(aes(color = like.), alpha = 1/2) + ggtitle("Danceability of Acoustic Songs")
ggplot(aes(x=instrumentalness, y=danceability), data = songs) + geom_point(aes(color = like.), alpha = 1/2) + ggtitle("Danceability of Instrumental Songs")
ggplot(aes(x=tempo, y = danceability), data = songs) + geom_point(aes(color = like.), alpha = 1/2) + ggtitle("Tempo vs. Danceability")
ggplot(aes(x=acousticness, y=liveness), data = songs) + geom_point(aes(color = like.), alpha = 1/2) + ggtitle("Relationship Between Acousticness and Probaiblity of Liveness")
```

To model the data, a decision tree can be used as a first pass.

```{r echo=FALSE, Model_Decision_Tree}

## 80% of the sample size
smp_size <- floor(0.80 * nrow(songs))

## set the seed to make your partition reproductible
set.seed(123)
train_ind <- sample(seq_len(nrow(songs)), size = smp_size)

train <- songs[train_ind, ]
test <- songs[-train_ind, ]

fit <- rpart(like. ~ acousticness + danceability + duration_ms + energy + instrumentalness + key + liveness + loudness + speechiness + tempo + valence,
               data=train,
               method="class")

fancyRpartPlot(fit)
Prediction <- predict(fit, test, type = "class")
predictions <- data.frame(test, prediction = Prediction)
paste0(round(100*with(predictions, mean(like.==prediction))), "%")
```
```{r echo=FALSE, Model_Random_Forest}
set.seed(415)
fit_rf <- randomForest(as.factor(like.) ~ acousticness + danceability + duration_ms + energy + instrumentalness + key + liveness + loudness + speechiness + tempo + valence,
                      data=train, 
                      importance=TRUE, 
                      ntree=2000)
varImpPlot(fit_rf)

Prediction_rf <- predict(fit_rf, test)
predictions_rf <- data.frame(test, prediction = Prediction_rf)
paste0(round(100*with(predictions_rf, mean(like.==prediction))), "%")
```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?
Songs I like typically have lower energy and loudness when compared to songs I do not like. Additionally, when danceability is coupled with tempo, songs I like appear to have lower danceability than songs I do not like.

### Were there any interesting or surprising interactions between features?
The interaction between tempo and danceability between songs I like and dislike is interesting because there was little difference when looking at danceability by itself. Only when splitting up danceability by tempo can I see the songs I like have lower danceabilty. This is because danceability and tempo have a quadratic relationship.

### Discuss the strengths and limitations of your model.
Random forests and decision trees have the benefit of being robust to outliers and they are able to find non-linear relationships within features. Individual trees suffer from overfitting if they are unconstrained.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
p1 <- ggplot(aes(x=like., y=acousticness), data = songs) + geom_boxplot() + ylab("Acousticness") + xlab("Preferred?")
p2 <- ggplot(aes(x=like., y=energy), data = songs) + geom_boxplot() + ylab("Energy") + xlab("Preferred?")
p3 <- ggplot(aes(x=like., y=instrumentalness), data = songs) + geom_boxplot() + ylab("Instrumentalness") + xlab("Preferred?")

grid.arrange(p1, p2, p3, ncol = 3)
```

### Description One
The plot above depicts the three song features that are most dissimilar between songs I prefer and songs I do not prefer. The box plots also show the wide range of values as seen by the many outliers. 

### Plot Two
```{r echo=FALSE, warning=FALSE, message=FALSE, Plot_Two}
ggplot(aes(x=loudness, y=energy), data = subset(songs, songs$loudness < quantile(songs$loudness, .99) & songs$loudness > quantile(songs$loudness, .01))) + geom_point(aes(color = like.), alpha = 1/2) + ggtitle("Relationship Between Loudness and Energy") + geom_smooth() + ylab("Energy") + xlab("Loudness (dB)") + labs(color = "Preferred?")
```

### Description Two
As the data was explored, it was found that energy and loudness have a positive correlation, as shown in the plot above. Furthermore, preferred songs tend to have slightly less energy and loudness than songs which are not preferred. 

### Plot Three
```{r echo=FALSE, Plot_Three}
fancyRpartPlot(fit)
```

### Description Three
Plot Three shows a visual representation of the decision tree model that was fitted to the data. By using instrumentalness, energy, speechiness, and acoustiness of the songs, the model was able to classify with 69% accuracy which songs were preferred.

------

# Reflection
The data exploration showed that preferred songs tend to show more instrumentalness, lower energy, and more acousticness than songs that were not preferred. With this knowledge, the random forest model that was created was able to improve song preferrence by 50% over the base case. Future extended modeling can be done with different machine learning techniques such as boosting to extend song recommendations.
